Analysis of the kinematics of punches in karate and their recognition using an artificial neural network

### 2021-04-15 mdpi/article.tex
Add files via upload
fix issues

\section{Materials and Methods}
\subsection{Participants}
One healthy participant (n=1), wight = 100 kg, height = 190 cm with experience in Karate took part in the study. Ethical approval was granted by the Human Research Ethics Commitee at Financial University at Government of the Russian Federation.
\subsection{Materials}
One 6DOF Inertial Measurement Unit (IMU) T-Wristband with properties (sampling rate 50 Hz, +- 16 g linear acceleration, +- 2000 deg/s gyroscope) was used in this study. T-wristband bracelet with 252x19x12 basic dimensions, 47x18x12 host dimension, 30 g weight with custom firmware used.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/IMU.jpg}
\caption{IMU, micro-controller and Bluetooth module}
\end{figure}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/hand_imu.jpg}
\caption{Box with IMU attached under bandage on wrist}
\end{figure}

@@

To record session we use Xiaomi Redmi 7 camera with 1980x1080 resolution 30 fps. Video analysis was used to labeling ground truth punches. To record data we use Bluetooth Serial Terminal Android application.
\subsection{Methods}
Data collection session consist of the participant performing 1912 punches to makivara.

Class of punches are:
\begin{enumerate}
\item	Yun Tsuki (YT); 
\item	Mawashi Tsuki (MT);
\item	Age Tsuki (AT);
\item	Uraken (U);
\item	No Punch (NP).
\end{enumerate}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/punch_names.jpg}
\caption{Punch classes}
\end{figure}

@@

Train / validation random splitting was made with 10:1 propotional for each class.

@@

 Data preprocessing was conducted with python 3.7 packages: numpy, sklearn. Visualisation was made with matplotlib, Neural Net models build with tensorflow.keras 2.2.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/data_xyz.png}
\caption{Measurement raw data visualized for different punch classes}
\end{figure}

In experiments takes part 4 models:
\begin{enumerate}
\item	multylayer perceptron; 
\item	1 dimension convolution network from scratch;
\item	2 dimension convolution network with 2 layers;
\item	2 dimension convolution network with 4 layers.
\end{enumerate}

@@

where $N_T$ - number of true classified punch,  
$N$ - total number of punches.\\
Precision, recall and F1-score were used as a classification metrics for single classes:
\begin{equation}
P = \frac{N_TP}{N_TP + N_FP}
\end{equation}
\begin{equation}
R = \frac{N_TP}{N_TP + N_FN}
\end{equation}
\begin{equation}
F1 = 2 \cdot \frac{P \cdot R}{P + R}
\end{equation}
where $N_TP$ - number of true positive classified punch,  
$N_FP$ - number of false positive classified punches, $P$ - precision, $R$ - recall, $F1$ - F1 - score.\\
Models was trained using PC with Ubuntu 18.04 LTS, Intel(E) Core(TM) i7-6950x CPU, 64 GB RAM, GTX 1080ti 8 GB GPU. Total time about 4 hours.

@@

\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

@@

\subsection{Simple multy layer perceptron}
Multy layer perceptron consist of 5 sequential layers with hidden size (450, 1024, 256, 128, 5), batch normalization, sigmoid and relu activations.
100 epochs training history and confusion matrix are on figure.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/mlp_train_log.png}
\caption{MLP training history}
\end{figure}

Classification metrics are on the table 1.

\begin{specialtable}[H] 
\caption{MLP classification metrics.}
\label{tab1}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.92		& 0.96		& 0.94\\
MT		& 0.96		& 0.90		& 0.93\\
AT		& 0.93		& 0.85		& 0.89\\
U		& 0.92		& 0.98		& 0.95\\
NP		& 0.85		& 0.87		& 0.86\\
\bottomrule
\end{tabular}
\end{specialtable}

Confusion matrix is on figure 7.

On training history we see difference between train and validation accuracy and small loss.
This means, that linear model is overfitting.
To avoid this we try more complex model - 1D convolution network.

@@

\subsection{1D Convolution Network}
1-dimension Convolution Network consist of 3 separate layers with 64 kernels each and relu activations. 
Optimizer is Adam, learning rate = 2e-3 and batch size 64.
100 epochs training history and confusion matrix are on figure.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/1conv_train_log.png}
\caption{1D CNN training history}
\end{figure}

Classification metrics are on the table 2.

\begin{specialtable}[H] 
\caption{1D CNN classification metrics\label{tab2}}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.41		& 0.91		& 0.56 \\
MT		& 0.75		& 0.06		& 0.11 \\
AT		& 0.70		& 0.79		& 0.74 \\
U		& 0.91		& 0.63		& 0.74 \\
NP		& 0.83		& 0.79		& 0.81 \\
\bottomrule
\end{tabular}
\end{specialtable}

On training history we see small train accuracy, unstable validation accuracy and big loss.
This means, that 1D conv model is unsuitable for punch classification.
So, we try the model with 2D convolution layers.

@@

\subsection{2D Convolution Network}
2-dimension Convolution Network consist of 2 layers, that inputs are both x,y and y,z axis. Layers has  72 and 88 kernels, size (2, 52), batch normalization and relu activations. 
Optimizer is Adam, learning rate = 2e-3 and batch size 64.
100 epochs training history and confusion matrix are on figure.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/2conv_train_log.png}
\caption{2D CNN training history}
\end{figure}

Classification metrics are on the table 3.

\begin{specialtable}[H] 
\caption{2D CNN classification metrics.\label{tab3}}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	&\textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.87		& 1.00		& 0.93 \\
MT		& 0.85		& 0.96		& 0.90 \\
AT		& 0.97		& 0.85		& 0.90 \\
U		& 0.95		& 0.85		& 0.90 \\
NP		& 0.97		& 0.87		& 0.92 \\
\bottomrule
\end{tabular}
\end{specialtable}

On training history we see much better train and validation accuracy, small loss.
But can we try deeper model with 4 2D convolution layers?

@@

\subsection{2D Convolution Network with 4 layers}

2-dimension Convolution Network consist of 4 layers, that inputs are both x,y and y,z axis. First layer has 32 kernels with size (2, 32), other layers has 64, 64 and 96 (2, 2) size kernels. Batch normalization and relu activations also used. 
Optimizer is Adam, learning rate = 2e-3 and batch size 64.
100 epochs training history and confusion matrix are on figure.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/4conv_train_log.png}
\caption{2D 4 layer CNN training history}
\end{figure}

Classification metrics are on the table 4.
% The MDPI table float is called specialtable
\begin{specialtable}[H] 
\caption{2D 4 layer CNN classification metrics.\label{tab4}}
%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.87		& 1.00		& 0.93 \\
MT		& 0.85		& 0.96		& 0.90 \\
AT		& 0.97		& 0.85		& 0.90 \\
U		& 0.95		& 0.85		& 0.90 \\
NP		& 0.97		& 0.87		& 0.92 \\
\bottomrule
\end{tabular}
\end{specialtable}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/4conv_confus.png}
\caption{2D 4 layer CNN confusion matrix}
\end{figure}

On training history we see much better train and validation accuracy, small loss.

\section{Discussion}
Multilayer perceptron is a simple model and good baseline. Best F1 score is 0.95 for U-punch, worst is 0.86 for N-punch. Difference between train and validation accuracy is rusult of model overfiiting.
Increasing number of training samples will solve this problem.\\
1D convolution model from [] work works very bad and not suitable for punch class prediction. Train accuracy is about 0.8, but validation accuracy only 0.65 and very unstable. Worst F1 score is 0.11 for MT-punch, best F1 is 0.81 for No-punch class. Loss after 100 epochs training is only about 0.3.\\
As we proposed, 2D convolution model with 2 conv layers works better. Metrics are like on mlp: best F1 is 0.93 for YT-punch and worst is 0.90 for U-punch class. A little gap between training and validation accuracy curves is tell about some overfitting, so we tested deeper conv model with 4 layers.\\
2D convolution model with 4 conv layer shows best result: 0.97 validation accuracy. Best F1-score 0.99 for YT-punch class, worst 0.90 for AT-punch class.\\
Comparing with MLP, we achieved better classification metrics and shift invariant model, based on 2D convolution.

@@

>>>>

\section{Materials and Methods}

\subsection{Participants}
The study involved sixteen healthy participants (n=16), 12 men, 4 women, aged 22±3 years, weighing = 70±14 kg, height = 165±21 cm with 3-7 years of experience in karate. Ethical approval was granted by the Human Research 76 Ethics Commitee at Financial University under the Government of the Russian Federation.

\subsection{Materials}
The design of the experiment can be seen in Figure 1.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/experiment_design.png}
\caption{Design of the experiment}
\end{figure}

On the wrists of the athletes were fixed devices, which included a microcontroller, IMUs, and Bluetooth modules. Athletes punched in shadow fight mode. The IMUs (accelerometer and gyroscope) data was initially transmitted via the Bluetooth channel to the android device. On the android device, the data was saved as files for each type of punch. This data was then processed on the computer. In order to label and identify each punch for develop models of the artificial network, video recordings of the experiments were made.
The data acquisition device (Figure 2) was a 50x20x10 mm box containing three modules – the microcontroller stm32f103 [9], IMU MPU9250 [10], and Bluetooth module HC-05 with a BC417143 chip [*].

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/IMU.png}
\caption{Data acquisition device}
\end{figure}

Figure 3 shows that the IMU device was attached to the athlete's wrist with boxing bandages. Figure 3 also shows the directions of the acceleration axes and the angular velocity of the gyroscope.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/bandage.png}
\caption{Measuring module attached to the arm with a boxing hand wrap}
\end{figure}

@@

To record session we use Xiaomi Redmi 7 camera with 1980x1080 resolution 30 fps. Video analysis was used to labeling ground truth punches. To record data we use Bluetooth Serial Terminal Android application.

Data collection session consist of the participant performing 1912 punches in shadow fight mode

Class of punches are:

\begin{enumerate}
\item	Yun Tsuki (YT);
\item	Mawashi Tsuki (MT);
\item	Age Tsuki (AT);
\item	Uraken (U);
\item	No Punch (NP).
\end{enumerate}

In Figures 4-7, the red arrow shows the approximate trajectories of the punches

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/YT.png}
\caption{Yun Tsuki (YT)}
\end{figure}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/MT.png}
\caption{Mawashi Tsuki (MT)}
\end{figure}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/AT.png}
\caption{Age Tsuki (AT)}
\end{figure}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/U.png}
\caption{Uraken (U}
\end{figure}

\subsection{Methods}

@@

Train / validation random splitting was made with 10:1 propotional for each class. Histogram of classes samples distribution on the Figure 8.

@@

 Data preprocessing was conducted with python 3.7 packages: numpy, sklearn. Visualisation was made with matplotlib, Neural Net models build with tensorflow.keras 2.2. Raw data for different punch classes visualized on Figure 9.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/data_xyz.png}
\caption{Measurement raw data for different punch classes}
\end{figure}

In experiments takes part 4 models:
\begin{enumerate}
\item	multylayer perceptron;
\item	1 dimension convolution network from scratch;
\item	2 dimension convolution network with 2 layers;
\item	2 dimension convolution network with 3 layers.
\end{enumerate}

@@

where $N_T$ - number of true classified punch,
$N$ - total number of punches.\\
Precision, recall and F1-score were used as a classification metrics for single classes:
\begin{equation}
P = \frac{N_{TP}}{N_{TP} + N_{FP}}
\end{equation}
\begin{equation}
R = \frac{N_{TP}}{N_{TP} + N_{FN}}
\end{equation}
\begin{equation}
F1 = 2 \cdot \frac{P \cdot R}{P + R}
\end{equation}
where $N_TP$ - number of true positive classified punch,
$N_FP$ - number of false positive classified punches, $P$ - precision, $R$ - recall, $F1$ - F1 - score.\\
Models was trained using PC with Ubuntu 18.04 LTS, Intel(E) Core(TM) i7-6950x CPU, 64 GB RAM, GTX 1080ti 8 GB GPU. Total time about 4 hours.

@@

\section{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Simple multy layer perceptron}
Multy layer perceptron consist of 5 sequential layers with hidden size (450, 1024, 256, 128, 5), batch normalization, sigmoid and relu activations. After 100 epochs training we have 0.99 training and 0.87 validation accuracy. Classification metrics are on the Table 1.

\begin{specialtable}[H]
\caption{MLP classification metrics.}
\label{tab1}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.87		& 0.93		& 0.91\\
MT		& 0.97		& 0.90		& 0.93\\
AT		& 0.83		& 0.90		& 0.87\\
U		& 0.92		& 0.96		& 0.94\\
NP		& 0.88		& 0.79		& 0.83\\
\bottomrule
\end{tabular}
\end{specialtable}

Confusion matrix is on Figure 10.

We see difference between train and validation accuracy, this means, that linear model is overfitting.
To avoid this we try more complicated model - 1D convolution network.

@@

\subsection{1D Convolution Network}
1-dimension Convolution Network consist of 3 separate layers with 64 kernels each and relu activations.
Optimizer is Adam, learning rate = 2e-3 and batch size 64.
After 100 epochs training we have 0.81 training and 0.62 validation accuracy. Confusion matrix is on Figure 11. Classification metrics are on the Table 2.

\begin{specialtable}[H]
\caption{1D CNN classification metrics\label{tab2}}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.64		& 0.85		& 0.73 \\
MT		& 0.91		& 0.63		& 0.75 \\
AT		& 0.86		& 0.75		& 0.81 \\
U		& 0.85		& 0.98		& 0.91 \\
NP		& 0.86		& 0.79		& 0.82 \\
\bottomrule
\end{tabular}
\end{specialtable}

During training process we observe small train accuracy, unstable validation accuracy and big loss. This means, that 1D convolution model is unsuitable for punch classification.
So, we try to use feature combination and model with 2D convolution layers.

@@

\subsection{2D Convolution Network}
2-dimension Convolution Network consist of 2 layers, that inputs are both x,y and y,z axis. Layers has  72 and 88 kernels, size (2, 52), batch normalization and relu activations.
Optimizer is Adam, learning rate = 2e-3 and batch size 64.
After 100 epochs training we have 0.97 training and 0.93 validation accuracy. Confusion matrix is on the Figure 12. Classification metrics are on the Table 3.

\begin{specialtable}[H]
\caption{2D CNN classification metrics.\label{tab3}}
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	&\textbf{precision}	& \textbf{recall}	& \textbf{F1-score}\\
\midrule
YT		& 0.81		& 1.00		& 0.89 \\
MT		& 1.00		& 0.94		& 0.97 \\
AT		& 0.90		& 0.85		& 0.88 \\
U		& 0.96		& 0.98		& 0.97 \\
NP		& 1.00		& 0.81		& 0.90 \\
\bottomrule
\end{tabular}
\end{specialtable}

We see much better train and validation accuracy, so we try deeper model with 3 2D convolution layers.

@@

\subsection{2D Convolution Network with 3 layers}

2-dimension Convolution Network consist of 3 layers, that inputs are both x,y and y,z axis. First layer has 32 kernels with size (2, 32), other layers has 64, 64 and 96 (2, 2) size kernels. Batch normalization and relu activations also used.
Optimizer is Adam, learning rate = 2e-3 and batch size 64. After 100 epochs training we have 0.98 training and 0.96 validation accuracy. Confusion matrix is on the Figure 13. Classification metrics are on the Table 4.
% The MDPI table float is called specialtable
\begin{specialtable}[H]
\caption{2D 3 layer CNN classification metrics.\label{tab4}}
%%% \tablesize{} %% You can specify the fontsize here, e.g., \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{cccc}
\toprule
\textbf{Punch class}	& \textbf{precision}	& \textbf{recall}	& \textbf{F1-score} \\
\midrule
YT		& 0.94		& 1.00		& 0.97 \\
MT		& 1.00		& 0.98		& 0.99 \\
AT		& 0.97		& 0.94		& 0.95 \\
U		& 0.96		& 1.00		& 0.98 \\
NP		& 1.00		& 0.92		& 0.96 \\
\bottomrule
\end{tabular}
\end{specialtable}

\begin{figure}[H]
\includegraphics[width=10.5 cm]{img/3conv_confus.png}
\caption{2D 3 layer CNN confusion matrix}
\end{figure}

On training history we see much better train and validation accuracy, small loss. We try more layers, but this do not improve classification metrics

@@

\section{Discussion}
Multilayer perceptron is a simple model and good baseline. Best F1 score is 0.95 for U-punch, worst is 0.86 for N-punch. Difference between train and validation accuracy is result of model overfiiting.
Increasing number of training samples will solve this problem.\\
1D convolution model from [] work works very bad and not suitable for punch class prediction. Train accuracy is about 0.8, but validation accuracy only 0.65 and very unstable. Worst F1 score is 0.11 for MT-punch, best F1 is 0.81 for No-punch class. Loss after 100 epochs training is only about 0.3.\\
As we proposed, 2D convolution model with 2 conv layers works better. Metrics are like on mlp: best F1 is 0.93 for YT-punch and worst is 0.90 for U-punch class. A little gap between training and validation accuracy curves is tell about some overfitting, so we tested deeper conv model with 4 layers.\\
2D convolution model with 3 conv layer shows best result: 0.97 validation accuracy. Best F1-score 0.99 for YT-punch class, worst 0.90 for AT-punch class.\\
Comparing with MLP, we achieved better classification metrics and shift invariant model, based on 2D convolution.

@@

\bibitem[stm32(2021)]{ref-url1}
STM32 overview. Available online: https://www.st.com/en/microcontrollers-microprocessors/stm32f103.html (accessed on April 9, 2021).

\bibitem[stm32(2021)]{ref-url2}
MPU-6000 product specification. Available online: https://invensense.tdk.com/wp-content/uploads/2015/02/MPU-6000-Datasheet1.pdf (accessed on April 9, 2021)
